Automated Dynamic Slice Backsolving

### 2014-03-24 docs/paper/fsuthesis/thesis-template/paper.tex
added cool figure for dynamic slice backsolving

TODO INSERT COOL FIGURE of custom data structure

@@

TODO
python specifics. Binarys, RE
how i implemented DTA
Pin's API

@@

There are two places we consider taint to propagate to and from, namely registers and memory locations. We also must record operations that transform tainted data, so that Stubble can backsolve the transformations later. We record this data in data structures derived from the vector, list, tuple and unordered\_map data structures from the C++ STL. The registers are represented as an enumerated array of tuples of (int, string). Registers start with tuples of (-1, ""), signifying that they are untainted. When registers become tainted, we insert values of (byteLocation, filename). Registers are altered as taint is overwritten or removed from the system, and since they only contain one value at a time, we mark the register as tainted by the specific byte in user input, using its location to refer to it. Memory is conceived of as an unordered\_map of ints to (int, string) tuples. This is similar to the registers, except the key is the memory location. Since there are a small set of registers it makes sense to design an array which contains all of them, but many diverse memory locations may become tainted throughout execution, so we instead use an unordered\_map. Memory is marked tainted or untainted similarly to registers. Lastly, we must record operations performed on tainted data. This is accomplished by representing user input as an array of lists. Each list entry is a string, marking one operation performed on the data. Stubble cross references each operation with the register and memory data structures to see which byte of user input should be updated with a transaction. After execution, a file of transactions for each byte of user input is printed, and the Python driver parses the operations recorded to backsolve new data values for each byte.

@@

@@

% undertainting examples, general DTA issues and how we solved.

@@

TODO we assume away shitty loops

%\section{Parallelization}

%Because Stubble works at the process level, allowing one run to finish before generating a test file for the next, concurrency seems implementable. However, because

@@

%We are leveraging Intel PIN to perform dynamic binary instrumentation. We are testing dummy programs with many paths that are easy to backsolve, similar to the example above. A C program which uses sequential switch statements to check user input against a secret, when compiled with tcc, which performs very little optimization, is such an example binary.

%There are allowable situations in programs which are outside the scope of this program. For instance, programs can loop forever [X]. These can be dealt with using reasonable heuristics, fuzzers tend to simply count loops and terminate after a set threshold is reached, and could be the subject of future research.

@@

%We implement dynamic taint analysis for files as input sources. We are basing our implementation of DTA on Jonathan Salwan\'s work [N]. Currently we instrument the read syscall, watching for all reads executed, then tag user input at the byte level and follow it as it moves throughout the binary. We instrument binaries at the assembly instruction level using Intel Pin, which dynamically disassembles each instruction. Each instruction is handled differently based upon the following criteria:

@@

%For example, move operations involving an untainted first operand and a tainted second operand generally spread taint, while those with a tainted first operand and an untainted second operand remove taint. Tainted branches trigger the backsolving technique. Various precautions, such as removing taint for an instruction like "xor eax, eax", must be taken. Lossy operations, like shifts, must be handled carefully to be backsolve-able, and checksums and cryptographic operations are not supported.

%As a demonstration we have run our fuzzer, Stubble, against various binaries which open a file and check specific bytes in a file for specific character values. If all of these checks pass, then the program outputs a "win" string, otherwise it outputs a "lose" string. This can be thought of conceptually as an oracle attack, and is similar to reverse engineering "crackme" exercises. Given no knowledge of the secret key, and only able to query an oracle, a successful attack must discover the secret within a reasonable number of guess [CITE?].

%After compiling with the Tiny C Compiler, which performs minimal optimization and doesn't add security measures, this binary and a default test file are given as input to Stubble, which will begin to dynamically generate run cases which will explore new paths within the binary. These test cases are fed back into the system, which continues to backsolve to find new ways to explore at branches, until there are no tainted bytes left for exploration.

%Professional fuzzing products, like Peach Fuzzer

%\section{Methodology}

%As of this writing we are moving towards fuzzing Linux utilities. Results for toy binaries are promising, with "crackme" type programs which ask for a secret being solved in approximately 2*n runs, where n is the length of the secret. There is, however, still some overtainting which must be accounted for.	
%\section{Evaluation}

@@

\section{Lessons Learned}

TODO RE'd in IDA Pro a lot
TODO learned about OS internals
TODO difficulties with implementation

>>>>

% TODO label this
\begin{tikzpicture}[node distance = 2cm, auto]
    % Place input bytes
    \node [block] (b0) {$b_0$};
    \node [block, right of=b0, node distance=2.2cm] (b1) {$b_1$};
    \node [block, right of=b1, node distance=2.2cm] (b2) {$b_2$};
    \node [block, right of=b2, node distance=2.2cm] (bs) {...};
    \node [block, right of=bs, node distance=2.2cm] (bn) {$b_n$};

	% Place transformation, flag, and branch nodes 
    \node [cloud, below of=b1] (t10) {$t_0$};
    \node [cloud, below of=t10, left of=t10] (t11) {$t_1$};
    \node [cloud, below of=t11, right of=t11] (flag) {flag};
    \node [cloud, below of=flag] (branch) {branch};
    \node [cloud, below of=bn] (flag1) {flag};
    \node [cloud, below of=flag1] (branch1) {branch};

    % Draw edges
    \path [line] (b1) -- (t10);
    \path [line] (t10) -- (t11);
    \path [line] (t11) -- (flag);
    \path [line] (flag) -- (branch);
    \path [line] (bn) -- (flag1);
    \path [line] (flag1) -- (branch1);
    %\path [line, dashed] (explored) -- (decide);

@@

% TODO Binarys, RE?
% TODO Pin's API more details?

@@

There are two places we consider taint to propagate to and from, namely registers and memory locations. We also must record operations that transform tainted data, so that Stubble can backsolve the transformations later. We record this data in data structures derived from the vector, list, tuple and unordered\_map data structures from the C++ STL. The registers are represented as an enumerated array of tuples of (int, string). Registers start with tuples of (-1, ""), signifying that they are untainted. When registers become tainted, we insert values of (byteLocation, filename). Registers are altered as taint is overwritten or removed from the system, and since they only contain one value at a time, we mark the register as tainted by the specific byte in user input, using its location to refer to it. The EFLAGS register is a special case, since certain operations set specific flags, it is possible for one operation to taint the carry flag and for another to taint the overflow flag. Thus, we treat each of the important bits of EFLAGS as separate entries, with their own tuple devoted to them.

Memory is conceived of as an unordered\_map of ints to (int, string) tuples. This is similar to the registers, except the key is the memory location. Since there are a small set of registers it makes sense to design an array which contains all of them, but many diverse memory locations may become tainted throughout execution, so we instead use an unordered\_map. Memory is marked tainted or untainted similarly to registers. Lastly, we must record operations performed on tainted data. This is accomplished by representing user input as an array of lists. Each list entry is a string, marking one operation performed on the data. Stubble cross references each operation with the register and memory data structures to see which byte of user input should be updated with a transaction. After execution, a file of transactions for each byte of user input is printed, and the Python driver parses the operations recorded to backsolve new data values for each byte.

@@

To test Stubble's ability to perform an oracle attack, we prepared C binaries which checked user input against a secret. After compiling with the Tiny C Compiler, which performs minimal optimization and doesn't add security measures like stack cookies, this binary and a default test file are given as input to Stubble.

@@

@@

% TODO we assume away shitty loops
%There are allowable situations in programs which are outside the scope of this program. For instance, programs can loop forever [X]. These can be dealt with using reasonable heuristics, fuzzers tend to simply count loops and terminate after a set threshold is reached, and could be the subject of future research.

% TODO undertainting examples, general DTA issues and how we solved.?

% Probably leave this out, parallelization didn't work out well b/c of input dependence
%\section{Parallelization}

@@

@@

@@

%For example, move operations involving an untainted first operand and a tainted second operand generally spread taint, while those with a tainted first operand and an untainted second operand remove taint. Tainted branches trigger the backsolving technique. Various precautions, such as removing taint for an instruction like "xor eax, eax", must be taken.

@@

% TODO not sure if to put this section in?
%\section{Experience and Lessons Learned}

%TODO RE'd in IDA Pro a lot
%TODO learned about OS internals
%TODO difficulties with implementation
